---
title: "Untitled"
author: "Moheth Muralidharan"
date: "10/27/2022"
output: pdf_document
---

#package installations
install.packages("ggpubr", repos = "https://cloud.r-project.org/", dependencies = TRUE)

##Libraries

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggpubr)
library(gridExtra)
```

##Importing the Datasets

```{r}
ow <- read.csv('/Users/tataurosetta/Downloads/Courses and Assignments/Computer Vizualization (IE 6600)/Homeworks/Homework 2/Datasets/DataOverweight.csv',
               na.strings="NA")
uw <- read.csv('/Users/tataurosetta/Downloads/Courses and Assignments/Computer Vizualization (IE 6600)/Homeworks/Homework 2/Datasets/DataUnderweight.csv',
              na.strings="NA")
stunt <- read.csv('/Users/tataurosetta/Downloads/Courses and Assignments/Computer Vizualization (IE 6600)/Homeworks/Homework 2/Datasets/DataStunted.csv',
                  na.strings="NA")
wasted <- read.csv('/Users/tataurosetta/Downloads/Courses and Assignments/Computer Vizualization (IE 6600)/Homeworks/Homework 2/Datasets/DataWasted.csv',
                   na.strings="NA")
fm <- read.csv('/Users/tataurosetta/Downloads/Courses and Assignments/Computer Vizualization (IE 6600)/Homeworks/Homework 2/Datasets/fm.csv',na.strings="NA")
dc_param <- read_excel("/Users/tataurosetta/Downloads/Courses and Assignments/Computer Vizualization (IE 6600)/Homeworks/Homework 2/Datasets/Data Country Parameters.xls",8)

```

## 1(a,b,c)

Perform the following steps to get the required table:
a) Join all 4 forms of malnutrition tables using joins on Country and Year in R
b) For each country, take the mean value for all the forms of malnutrition
c) Import the Data country parameters dataset and using joins, add WHO Region 
and Income group from the sheet 8 of the dataset.

#Code

```{r}
df1 <- ow %>% full_join (stunt,by=c("Country","Year"))
df2 <- df1%>% full_join (uw,by=c("Country","Year"))
hw_1 <- df2%>% full_join (wasted,by=c("Country","Year"))

colnames(hw_1) <- c('Country','Year','OverWeight','Stunted','UnderWeight',
                    'Wasted')

hw_1$OverWeight <- as.numeric(hw_1$OverWeight)
hw_1$Stunted <- as.numeric(hw_1$Stunted)
hw_1$UnderWeight <- as.numeric (hw_1$UnderWeight)
hw_1$Wasted <- as.numeric (hw_1$Wasted)

mut_hw1 <- hw_1%>%
  select (Country,OverWeight,Stunted,UnderWeight,Wasted)%>%
  group_by (Country)%>%
  mutate (AverageOverweightProportion=mean(OverWeight))%>%
  mutate (AverageStuntedProportion=mean(Stunted))%>%
  mutate (AverageUnderweightProportion=mean(UnderWeight))%>%
  mutate (AverageWastedProportion=mean(Wasted))

mut_hw1 <- as.data.frame (mut_hw1)

mut_hw1 <- mut_hw1 %>% 
  mutate_if (is.numeric, ~replace_na (.,mean (., na.rm = TRUE)))

mut_hw1 <- mut_hw1%>%
  select (Country,AverageOverweightProportion,AverageStuntedProportion,
         AverageUnderweightProportion,AverageWastedProportion)

mut_hw1 <- mut_hw1[!duplicated(mut_hw1), ]

temp_hw1 <- mut_hw1
rownames (temp_hw1) <- 1:nrow (temp_hw1)

dcparam_new <- data.frame (dc_param$DisplayValue,dc_param$WHO_REGION,
                         dc_param$`World Bank income group`)

colnames(dcparam_new)[1] <- "Country"

final_hw1 <- left_join (temp_hw1,dcparam_new, by='Country')
colnames (final_hw1)[6:7] <- c("WHO_REGION","World Bank Income group")

head (final_hw1,15)

```

#Conclusion

We are using full_join (outer join) to join the values of Overweight, Stunted,
Underweight, Wasted with respect to Country and Year. For each country, the
mean of each values is calculated. The WHO Region and Income group are also 
added.

## 2

Using the datasets provided, plot the graphs as shown below with titles and labels. (15 points)

View(final_hw1)

#Code
```{r}

hw_2 <- final_hw1

#overweight
hw_2_bar_ov <- hw_2 %>%
  select(WHO_REGION,`World Bank Income group`, AverageOverweightProportion) %>%
  group_by(WHO_REGION,`World Bank Income group`) %>%
  summarise(Count = round(sum(AverageOverweightProportion)/n()), .groups = 'drop') %>%
  arrange(desc(Count)) %>%
  drop_na()

gg_ov <-ggplot(hw_2_bar_ov, aes(fill=`World Bank Income group`, y=WHO_REGION, x=Count)) +
    geom_bar(position="dodge", stat="identity")+
  scale_fill_brewer(palette = 1)+
  labs(fill = 'Income levels of country', x = 'Percentage(%)')+
  geom_text(aes(label=Count), vjust=0.5,hjust=-0.5,position = position_dodge(width = 1))

gg_ov


#stunted
hw_2_bar_st <- hw_2 %>%
  select(WHO_REGION,`World Bank Income group`, AverageStuntedProportion) %>%
  group_by(WHO_REGION,`World Bank Income group`) %>%
  summarise(Count = round(sum(AverageStuntedProportion)/n()), .groups = 'drop') %>%
  arrange(desc(Count)) %>%
  drop_na()

gg_st <-ggplot(hw_2_bar_st, aes(fill=`World Bank Income group`, y=WHO_REGION, x=Count)) +
    geom_bar(position="dodge", stat="identity")+
  scale_fill_brewer(palette = 1)+
  labs(fill = 'Income levels of country', x = 'Percentage(%)')+
  geom_text(aes(label=Count), vjust=0.5,hjust=-0.5,position = position_dodge(width = 1))

gg_st


#underweight
hw_2_bar_ud <- hw_2 %>%
  select(WHO_REGION,`World Bank Income group`, AverageUnderweightProportion) %>%
  group_by(WHO_REGION,`World Bank Income group`) %>%
  summarise(Count = round(sum(AverageUnderweightProportion)/n()), .groups = 'drop') %>%
  arrange(desc(Count)) %>%
  drop_na()

gg_ud <-ggplot(hw_2_bar_ud, aes(fill=`World Bank Income group`, y=WHO_REGION, x=Count)) +
    geom_bar(position="dodge", stat="identity")+
  scale_fill_brewer(palette = 1)+
  labs(fill = 'Income levels of country', x = 'Percentage(%)')+
  geom_text(aes(label=Count), vjust=0.5,hjust=-0.5,position = position_dodge(width = 1))

gg_ud


#wasted
hw_2_bar_wt <- hw_2 %>%
  select(WHO_REGION,`World Bank Income group`, AverageWastedProportion) %>%
  group_by(WHO_REGION,`World Bank Income group`) %>%
  summarise(Count = round(sum(AverageWastedProportion)/n()), .groups = 'drop') %>%
  arrange(desc(Count)) %>%
  drop_na()

gg_wt <-ggplot(hw_2_bar_wt, aes(fill=`World Bank Income group`, y=WHO_REGION, x=Count)) +
    geom_bar(position="dodge", stat="identity")+
  scale_fill_brewer(palette = 1)+
  labs(fill = 'Income levels of country', x = 'Percentage(%)')+
  geom_text(aes(label=Count), vjust=0.5,hjust=-0.5,position = position_dodge(width = 1))

gg_wt



```





## 3

Using the datasets, fill in the correlation values between different forms of malnutrition in the table. 

#Code
```{r}
hw_3 <- hw_1
hw_3 <- hw_3 %>% 
  mutate_if (is.numeric, ~replace_na (.,mean (., na.rm = TRUE)))

cor (hw_3[,c('OverWeight','Stunted','UnderWeight','Wasted')])

```

#Conclusion

The correlation values are 1 when one variable is compared to the same variable.
Negative correlation means that it is correlated in the opposite direction.
If the correlation value is 0, it means that there is no correlation between 
those two variables.

## 4
Using the datasets, plot the following Violin and Box plots. (20 points)

#Box Plots
```{r}

hw_4 <- final_hw1

#Overweight Category
g_over <- ggplot(hw_4 , aes(x="",y=AverageOverweightProportion))+
  geom_boxplot()

g_over <- g_over + labs(x="  Overweight \n Category", 
       y="Value")

#Stunted Category
g_stunted <- ggplot(hw_4 , aes(x="",y=AverageStuntedProportion))+
  geom_boxplot()

g_stunted <- g_stunted + labs(x="  Stunted \n Category", 
       y="Value")

#Underweight Category
g_under <- ggplot(hw_4 , aes(x="",y=AverageUnderweightProportion))+
  geom_boxplot()

g_under <- g_under + labs(x="  Underweight \n Category", 
       y="Value")

#Wasted Category
g_wasted <- ggplot(hw_4 , aes(x="",y=AverageWastedProportion))+
  geom_boxplot()

g_wasted <- g_wasted + labs(x="  Wasted \n Category", 
       y="Value")

grid.arrange(g_over,
             g_stunted,
             g_under,
             g_wasted,
             nrow = 2, ncol = 2, heights=c(4,4),top = "Box plot for values of different forms of malnutrition")

```
```{r}

hw_4 <- final_hw1

#Overweight Category
g_over_v <- ggplot(hw_4 , aes(x="",y=AverageOverweightProportion))+
  geom_violin()

g_over_v <- g_over_v + labs(x="  Overweight \n Category", 
       y="Value")

#Stunted Category
g_stunted_v <- ggplot(hw_4 , aes(x="",y=AverageStuntedProportion))+
  geom_violin()

g_stunted_v <- g_stunted_v + labs(x="  Stunted \n Category", 
       y="Value")

#Underweight Category
g_under_v <- ggplot(hw_4 , aes(x="",y=AverageUnderweightProportion))+
  geom_violin()

g_under_v <- g_under_v + labs(x="  Underweight \n Category", 
       y="Value")

#Wasted Category
g_wasted_v <- ggplot(hw_4 , aes(x="",y=AverageWastedProportion))+
  geom_violin()

g_wasted_v <- g_wasted_v + labs(x="  Wasted \n Category", 
       y="Value")

grid.arrange(g_over_v,
             g_stunted_v,
             g_under_v,
             g_wasted_v,
             nrow = 2, ncol = 2, heights=c(4,4),top = "Violin plot for values of different forms of malnutrition")


```

## 5

Write a code to compute the number of farmers markets by state and arrange them
in descending order of number of farmers market.

#Code
```{r}
fm$FMID <- unique(fm$FMID)

state_fm <- fm%>%
  group_by(State)%>%
  summarise(Number_of_farmer_markets=n())%>%
  arrange(desc(Number_of_farmer_markets))
state_fm
```
#Conclusion

The number of farmer markets are arranged by state and in descending order.

##6

Write a code to compute the number of farmers market by cities in Massachusetts and 
display top five cities. 

#Code
```{r}
mass_fm <- fm%>%
  filter (State == 'Massachusetts', na.rm=TRUE)%>%
  group_by (city)%>%
  count (city,sort = TRUE)%>%
  drop_na()
head (mass_fm,5)
```

#Conclusion

The top five cities in the state of Massachusetts which have the highest number 
of farmer markets are displayed.

## 7

Generate the following table using pivot function. First column should contain 
the states.Second column should have Payment System. For Payment System consider
the columns, “Credit”, “WIC”, “WICcash”, and “SNAP” from the original farmers
market data. Third column should have the number of farmers market offering the payment services. 

#Code
```{r}
temp_fm <- fm %>%
  pivot_longer (cols=c('Credit','WIC','WICcash','SNAP'),
               names_to="Payment System",values_to="PS_values")

table_fm <- temp_fm%>%
  filter (PS_values=="Y")%>%
  group_by (State,`Payment System`)%>%
  summarise (Farmer_Markets=n())
table_fm
```

#Conclusion

The table is generated using the pivot_longer() function with the requirements
 accordingly.