---
title: "Visualization_part1"
output: html_document
---
```{r}
#depending on the timing we might work with the plotly library as well

library(ggplot2)
library(dplyr)
library(gridExtra)
# Use ggplot function to plot
# Using only ggplot function will create a blank canvas
 
head(mpg) #Table

head(mtcars)# data frame

ggplot()
```

```{r}
# Provide aesthetics to blank canvas - it does not show the data points unless its provided
# 
# Aesthetic mappings describe how variables in the data are mapped to visual properties (aesthetics) of geoms

ggplot(mpg, aes(x=cty,y=hwy))
```

```{r}
# In the first class we talked about the marks and attributes


# Generate canvas + add axis + plot data
# Geom provides layer to the graph - first lets check the geom by searching for it


ggplot(mpg, aes(x=cty,y=hwy))+geom_point()


ggplot(mpg, aes(x=cty,y=hwy))+geom_point(alpha = 1/10)


ggplot(mpg, aes(x=cty,y=hwy))+geom_point(shape = "+")


ggplot(mpg, aes(x=cty,y=hwy))+geom_point(shape = "+", colour = "red", size = 3)

ggplot(mpg, aes(x=cty,y=hwy)) +
  geom_point(shape = 21, colour = "black", fill = "white", size = 5, stroke = 5) #stroke is for the edges

```

```{r}
# Map additional dimension to aesthetics. In this case color
ggplot(mpg, aes(x=cty,y=hwy, color=class))+geom_point()

```

```{r}
# Map dimension as shape
ggplot(mpg, aes(x=cty,y=hwy, shape=drv))+geom_point()
```

```{r}
# Map dimension as size
ggplot(mpg, aes(x=cty,y=hwy, size=displ))+geom_point()
```

```{r}
# Mapping multi dimensions 
# 
# imagine we want to show the relation between city miles per gallon and highway miles per gallon
# for each the type of drive train (drv), now we want to see engine displacement, in litres as well
# we can add the fourth dimension
# 
# What can we see? we can show three infromation using one plot - for example for each
# green (front wheel) we can see as the milage per cty and highway increases the engine displacement decreases
ggplot(mpg, aes(x=cty,y=hwy, size=displ, color=drv))+geom_point()
```

```{r}
# Limitations of mapping according to type of variable
ggplot(mpg, aes(x=cty,y=hwy))+geom_point()
```


# Bar Plot
```{r}
df <- data.frame(dose=c("D0.5", "D1", "D2", "D2"),
                len=c(4.2, 10, 29.5, 15))


ggplot(data=df, aes(x=dose)) +
  geom_bar() #by default is stat = count

# First thing to see is that we are not providing the y-values

ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")

# If you explicitly say stat = "identity" in geom_bar(), you're telling 
# ggplot2 to skip the aggregation and that you'll provide the y values
```

```{r}
ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")+coord_flip()
```

```{r}
library(gridExtra)
p1<- ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")
p2<- ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")+coord_flip()
grid.arrange(p1,p2,ncol=2)
```

#for numerical variables (cylenders are numerical)

```{r }

ggplot(mtcars, aes(x=(cyl)))+
  geom_bar(stat="bin")

# What if we use a variable with factor levels ? 

# ggplot(mtcars, aes(x=(factor(cyl))))+
#   geom_bar(stat="bin")

```

# Stacked bar chart
```{r}
df2 <- data.frame(supp=rep(c("VC", "OJ"), each=3), #function rep dopes repeat the values for # of times specified
                dose=rep(c("D0.5", "D1", "D2"),2),
                len=c(6.8, 15, 33, 4.2, 10, 29.5))

p1<- ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity")
p2<- p1+coord_flip()
grid.arrange(p1,p2,ncol=2)
```

# Generating stacked bar chart as percentages
```{r}

head(mtcars)

library(scales)
# first we want to create a grouped Bar Graph for Comparison
# then we are creating a stacked bar graph for comparison
# which one do you think does a better work in representing the information? 

p1<-ggplot(mtcars, aes(factor(cyl), mpg, fill=factor(gear) )) +
  geom_bar(stat="identity", position="dodge") +coord_flip()+
  ggtitle("100 % Stacked Bar Graph for Comparison")+
  scale_y_continuous(labels = percent_format()) 


p2<-ggplot(mtcars, aes(factor(cyl), mpg, fill=factor(gear) )) +
  geom_bar(stat="identity", position="fill" ) +
  ggtitle("100 % Stacked Column Graph for Comparison")+
  scale_y_continuous(labels = percent_format())

grid.arrange(p1, p2,ncol = 2)
```


# Grouped bar chart
```{r}
p1<- ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity", position="dodge")
p2<- p1+coord_flip()
grid.arrange(p1,p2,ncol=2)
```

# Point to remember in bar chart
```{r}
# Make sure one axis is categorical
ggplot(mtcars, aes(x = cyl)) + geom_bar()
```

# Highlighting - lets color a specific condition
```{r}
library(magrittr)
midwest %>% 
  mutate(highlight_flag = ifelse(state == 'IL', T, F)) %>%  #lets create a new feature (T/F)
  ggplot(aes(x = state, fill = highlight_flag)) +
    geom_bar()

# Question: Can you plot this in a easier way?


ggplot(midwest, aes(x = state, fill = state == 'IL')) +
    geom_bar()
```

# Ordering categories 
```{r}
p1 <- ggplot(mtcars, aes(x = row.names(mtcars), mpg)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        ggtitle("Fig. A: Default rotated x-axis")

# order bars
# point here is when you are reordering the row.names it should be applied the whole dataset
# and then passed to the plot function - you are doing this by reordering the row names as an aesthetics
p2 <- ggplot(mtcars, aes(x = reorder(row.names(mtcars), mpg), y = mpg)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        ggtitle("Fig. B: Rotated ordered x-axis")

grid.arrange(p1, p2, ncol = 2)
```


# Text annotation
```{r}
# It is highly recommeded to use text annotation in the bar graph to make it easier for 
# your audience to follow or understand the graph - we can do this using geom_text() 

# Outside bars
p1<-ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=len), vjust=-.3, size=3.5)


# What if we decrease the vjust
#  
# p1<-ggplot(data=df, aes(x=dose, y=len)) +
#   geom_bar(stat="identity")+
#   geom_text(aes(label=len), vjust=-10.3, size=3.5)

# Inside bars
p2<-ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=len), vjust=1.6, color="white", size=3.5)
grid.arrange(p1,p2, ncol=2)
```
# Text annotation for cumulative sum in stacked bar chart

```{r}
library(plyr)
df_sorted <- arrange(df2, dose, supp)  # orders the rows of a data frame by the values of selected columns.
df_cumsum <- ddply(df_sorted, "dose",     #create a variable (label_ypos) for cumulative sum of (len) over variable (dose)
                   transform, label_ypos=cumsum(len))

ggplot(data=df_cumsum, aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity")+
  geom_text(aes(y=label_ypos, label=len), vjust=1.6, 
            color="white", size=3.5)
```

# Text annotation for cumulative sum in grouped bar chart


```{r}
ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=len), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)


ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity", position='dodge')+
  geom_text(aes(label=len), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)
```

# Augmenting additional visual elements

```{r}
# Adding reference line - sometimes we want to add a reference line to emphesize on a hidden knowledge/facts - we can use the refrence line for ex

  ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")+
  geom_hline(yintercept=20, linetype="dashed")
```

```{r}
  ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")+
  geom_hline(yintercept=mean(df$len), linetype="dashed")
```

```{r}
# Adding text
ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")+
  geom_hline(yintercept=mean(df$len), linetype="dashed")+
 geom_text(x=2, y=16, label="Mean Value")
```

```{r}
ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")+
  geom_hline(yintercept=mean(df$len), linetype="dashed")+
 annotate(geom="text",x=2, y=16, label="Mean Value")
```

# Alternate syntax
```{r}
## you could pass the dataset in the geom layer

df <- data.frame(dose=c("D0.5", "D1", "D2"),
                len=c(4.2, 10, 29.5))

ggplot() + 
  geom_bar(data=df, aes(x=dose, y=len),stat="identity")
```

```{r}
df <- data.frame(dose=c("D0.5", "D1", "D2"),
                len=c(4.2, 10, 29.5))

ggplot(data=df) +
  geom_bar( aes(x=dose, y=len),stat="identity")
```



# Heat Maps

```{r}
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5) # uniform distribution on the interval from min to max
 
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile()
```


# Heat Maps

```{r}
library(reshape2)
mydata <- mtcars[, c(1,3,4,5,6,7)]
cormat <- round(cor(mydata),2)
melted_cormat <- melt(cormat) # it change the wide to long format
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

```


# Data Distribution-Box Plot
```{r}
# The Effect of Vitamin C on Tooth Growth in Guinea Pigs

p1<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_boxplot()
p2<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_boxplot()+coord_flip() 

# The notch displays a confidence interval around the median which is normally based on the median +/- 1.58*IQR/sqrt(n). Notches are used to compare groups; if the notches of two boxes do not overlap, this is a strong evidence that the medians differ.

p3<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_boxplot(notch=TRUE) + theme(legend.position="bottom")

p4<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4)
grid.arrange(p1, p2, p3, p4,ncol = 2, heights=c(1,1))
```



```{r}
# Dont use fill+color=same variable for box plot

ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose))) + 
  geom_boxplot()

ggplot(ToothGrowth, aes(x=factor(dose), y=len, color=factor(dose))) + 
  geom_boxplot()

ggplot(ToothGrowth, aes(x=factor(dose), y=len, color=factor(dose), fill=factor(dose))) + 
  geom_boxplot()
```

# Box plot with mean points

```{r}

#The function stat_summary() can be used to add mean points to a box plot

ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_boxplot()+
  stat_summary(fun=mean, geom="point", size=2,color="red")


# You can also choose the item to display 
# 
ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_boxplot()+
  stat_summary(fun=mean, geom="point", size=2,color="red") +
  scale_x_discrete(limits=c("0.5", "2"))


```

# Box plot with multiple groups

```{r}
# Change box plot colors by groups

bp <- ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=supp))

bp + geom_boxplot()


# add  some distance between the boxplots of the same group
bp + geom_boxplot(position=position_dodge(1))

bp + geom_boxplot(position=position_dodge(1)) +
  scale_fill_brewer(palette="Blues") + theme_classic()
# Discrete colors
bp + geom_boxplot(position=position_dodge(1)) + 
  scale_fill_brewer(palette="Dark2") + theme_minimal()
# Gradient colors
bp + geom_boxplot(position=position_dodge(1)) + 
  scale_fill_brewer(palette="RdBu") + theme_minimal()



```
# Data Distribution-Violin Plots
```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_violin()
```

```{r}
p1<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_violin()


p2<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_violin()+coord_flip()
grid.arrange(p1, p2,ncol = 2, heights=c(1,1))

```

#Add a boxlot within the violin plot

```{r}
p1<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_violin()+geom_boxplot(width=0.1)

p2<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_violin()+coord_flip()+geom_boxplot(width=0.1)

grid.arrange(p1, p2,ncol = 2, heights=c(1,1))

```
```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len,fill=supp)) + 
  geom_violin()
```

# Data Distribution-Dot Plot
```{r}

# binaxis: The axis to bin along, "x" (default) or "y"
# 
# stackdir which direction to stack the dots. "up" (default), "down", "center",
# "centerwhole" (centered, but with dots aligned)

ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)

# 0.2 : degree of jitter in x direction

ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1) + 
  geom_jitter(shape=2, position=position_jitter(0.2))
```


# Change dots shape and fonts

```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_dotplot(binaxis='y', stackdir='center',
               stackratio=1.5, dotsize=1.2)
```
# Direction to stack the dots

```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_dotplot(binaxis='y', stackratio=1.5, dotsize=1.2)


ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_dotplot(binaxis='y', stackdir='down', dotsize=1)



```
# Adding statistcal  summary to the point (Mean, Median)

```{r}
p1<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_dotplot(binaxis='y', stackdir='center')+
  stat_summary(fun = mean, geom="point", shape=18,size=3, color="red")

p2<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_dotplot(binaxis='y', stackdir='center',
               stackratio=1.5, dotsize=1.2)+
  stat_summary(fun=mean, geom="point", shape=18,size=3, color="red")

p3<-ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_dotplot(binaxis='y', stackdir='center',
               stackratio=1.5, dotsize=1.2)+
  coord_flip()+
  stat_summary(fun=mean, geom="point", shape=18,size=3, color="red")

grid.arrange(p1, p2, p3,ncol = 3, heights=c(1,1))
```
```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=supp)) +
  geom_dotplot(binaxis='y', stackdir='center')
# use position=dodge
```

#Adding points to the violin plot

```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  geom_violin()+geom_dotplot(binaxis='y', stackdir='center', dotsize=1)
```


```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=supp)) + 
  geom_violin()+
  geom_dotplot(binaxis='y', stackdir='center', position="dodge")
```

# Untill here

## Single Variable - Histogram
```{r}
ggplot(mpg, aes(x = cty)) +
        geom_histogram()
```

```{r}
ggplot(mpg, aes(x = cty)) +
        geom_histogram(binwidth = 1, color = "grey30", fill = "white")
```

# Adding the mean line to the histogram
```{r}
ggplot(mpg, aes(x = cty)) +
        geom_histogram(binwidth = 1, color = "grey30", fill = "white")+
  geom_vline(xintercept = mean(mpg$cty), color = "red", linetype = "dashed", size=2)
```
# Lets add a third variable

```{r}

temp<-select(mpg, manufacturer, cty)
temp<-filter(temp, manufacturer=="audi"|manufacturer=="dodge")
ggplot(temp, aes(x = cty, fill=manufacturer))+
  geom_histogram(alpha = .5) +
  theme_update(text = element_text(size=16))
```

# show the mean value - good for undrestanding the distribution

```{r}
temp<-select(mpg, manufacturer, cty)
temp<-filter(temp, manufacturer=="audi"|manufacturer=="dodge")
compare_mean<-summarise(group_by(temp,manufacturer),Mean=mean(cty))

ggplot(temp, aes(x = cty, fill=manufacturer))+
  geom_histogram()+ 
  geom_vline(xintercept = compare_mean$Mean,color = "red", linetype = "dashed", size=2)+
  theme_update(text = element_text(size=16))
 
```

# Density plot - fill color
```{r}
ggplot(mpg, aes(x = cty)) +
       geom_density(fill="steelblue")
```

# Density plot - add a third variable

```{r}
temp<-select(mpg, manufacturer, cty)
temp<-filter(temp, manufacturer=="audi"|manufacturer=="dodge")

ggplot(temp, aes(x = cty, fill=manufacturer))+
  geom_density(alpha = .5) 
```

# Two dimension histogram


```{r}
ggplot(faithful, aes(x=eruptions, y=waiting)) +geom_bin2d()
# Each bin corresponds to series of (x and y)

```
# Density 2d with different geoms 


```{r}
p<- ggplot(faithful, aes(x=eruptions, y=waiting)) +
  geom_point()

p1 <- p + stat_density2d()

p3<- p +  stat_density2d(aes(color  = ..level..)) # We pass a Y variable which is generated using a formula as levels

p3<- p +  stat_density2d(aes(fill = ..level..), geom="polygon")

p4<- p +  stat_density2d(aes(fill = ..density..), geom = "raster", contour = FALSE) # We pass a Y variable which is generated using a density function

# To draw contours, you need breaks, whereas tiles/rasters are continuous if passed ..density..
#  
p5<- p +  stat_density2d(aes(alpha = ..density..), geom = "tile", contour = FALSE)

grid.arrange(p1, p2, p3, p4, p5, ncol = 2, nrow =3 )

```

## Simple Scatter Plots
### Uses: Relationship between variables


```{r}
sample_data<-mpg
ggplot(sample_data,aes(cty,hwy))+geom_point()
```

```{r}

ggplot(sample_data,aes(cty,hwy,color=factor(cyl),shape=factor(drv),size=displ))+
  geom_point()


# do you think this is a good way of visualization? 
```

```{r}
sample_data<-mpg
cor(sample_data$cty,sample_data$hwy)
cor(sample_data$hwy,sample_data$cty)

fit1<-lm(hwy~cty,sample_data)
fit2<-lm(cty~hwy,sample_data)
summary(fit1)
summary(fit2)
```

# draw patterns

```{r}

help("geom_smooth") 


ggplot(mpg, aes(x=cty, y=hwy)) + 
  geom_point()+ geom_smooth(method=lm, se=TRUE) # Aids the eye in seeing patterns in the presence of overplotting
```



```{r}
ggplot(mpg, aes(x=cty, y=hwy)) + 
  geom_point()+ geom_smooth()
```
 
# Draw relationship line per category (third variable) 
 
```{r}
ggplot(mpg, aes(x=cty, y=hwy, color=factor(cyl))) + 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)
```

# Draw relationship line (full range) per category (third variable) 


```{r}
ggplot(mpg, aes(x=cty, y=hwy, color=factor(cyl))) + 
  geom_point()+ 
  geom_smooth(method=lm, se=FALSE,fullrange=TRUE)
```

# Correlation plot

```{r}
library(corrplot)
M<-cor(mtcars)
head(round(M,2))
corrplot(M, method="circle")

```


```{r}
corrplot(M, method="pie")
```

```{r}
corrplot(M, method="color")
```


```{r}
corrplot(M, method="number")
```

```{r}
corrplot(M, type="upper")
```

# Ordering of the correlation matrix

```{r}
help("corrplot")
corrplot(M, type="upper", order="hclust")
corrplot(M, type="upper", order="alphabet")


```


## Scaling in Scatter Plot

```{r}
library(MASS)
ggplot(Animals, aes(x = body, y = brain)) + geom_point()
```

# logarithmic scaling and updating logticks

```{r}

p1<-ggplot(Animals, aes(x = body, y = brain)) + 
  geom_point() +
     scale_x_log10() +
     scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) 

p2<-ggplot(Animals, aes(x = body, y = brain)) + geom_point() +
     scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks()  

grid.arrange(p1, p2,ncol = 2)
```


# Time series
# Line plots


```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)

y<- sample(1:10, 100, replace=T)
x<-1:100
df<-data.frame(x,y)
ggplot(df, aes(x=x, y=y)) +
  geom_line()+
  ggtitle("x and y are continuous variables")
```

<!-- group="whatever" is a "dummy" grouping to override the default behavior, which (here) -->
<!-- is to group by dose and in general is to group by the x variable. in order to separately count the number of  -->
<!-- rows in each level of the x variable. For example, here, the default would be for -->
<!-- geom_line to return the number of rows with cut equal to "D0.5, "D1", "D2". we using group = 1 -->


```{r}
df2 <- data.frame(supp=rep(c("VC", "OJ"), each=3),
                dose=rep(c("0.5", "1", "2"),2),
                len=c(6.8, 15, 33, 4.2, 10, 29.5))

p1<-ggplot(subset(df2,df2$supp=="VC"), aes(x=dose, y=len, group=1)) +
  geom_line(color="red")+
  ggtitle("x is categorical and y is continuous")

p2<-ggplot(subset(df2,df2$supp=="OJ"), aes(x=dose, y=len,group=1)) +
  geom_line(color="orange")+
  ggtitle("x is categorical and y is continuous")

grid.arrange(p1, p2,ncol = 2)
```

# Grouping by a third variable

```{r}
ggplot(df2, aes(x=dose, y=len, group=supp, color=supp)) +
  geom_line()+
  ggtitle("x is categorical and y is continuous")
```

```{r}
ggplot(df2, aes(x=dose, y=len, group=supp)) +
  geom_line()+
  geom_point()+
  ggtitle("x is categorical and y is continuous")
```

#Converting x to continues  - shape changes as the distance

```{r}
ggplot(df2, aes(x=as.numeric(dose), y=len, group=supp, color=supp)) +
  geom_line()+
  geom_point()+
  ggtitle("x is continuous and y is continuous")
```


```{r}
ggplot(df2, aes(x=as.numeric(dose), y=len, group=supp, color=supp)) +
  geom_line(size=2)+ 
  geom_point(size=2)+
  ggtitle("x is continuous and y is continuous")
```

```{r}
ggplot(df2, aes(x=as.numeric(dose), y=len, group=supp, color=supp)) +
  geom_line()+ 
  geom_point(size=2)+
  ggtitle("x is continuous and y is continuous")
```

```{r}
ggplot(df2, aes(x=as.numeric(dose), y=len, group=supp, color=supp)) +
  geom_line(linetype="dashed")+ 
  geom_point(size=2)+
  ggtitle("x is continuous and y is continuous")
```

```{r}
ggplot(df2, aes(x=as.numeric(dose), y=len, group=supp, color=supp)) +
  geom_line(aes(linetype=supp))+
  geom_point(size=2)+
  ggtitle("x is continuous and y is continuous")
```

```{r}
ggplot(df2, aes(x=as.numeric(dose), y=len, group=supp, color=supp)) +
  geom_line(aes(linetype=supp))+ 
  geom_point(aes(shape=supp))+
  ggtitle("x is continuous and y is continuous")
```

