---
title: "Part 2"
output: html_document
---

library(ggplot2)
library(dplyr)
library(gridExtra)


## Stacked Area Charts
### Uses: comparison,part to whole

# For example imagine we want to compare the amount of each cut for each price (its easier to do this using area) 

```{r}
library(scales)
ggplot(diamonds, aes(x = price, fill=cut))+
  geom_area(stat="bin")+
  ggtitle("Stacked Area Graph for Comparison + Part to Whole")
```
# We could use this as proportion 

```{r}

# To verify lets get the proportion of each cut
data <- diamonds  %>%
  group_by(cut) %>%
  dplyr::summarise(n = sum(price)) %>%
  dplyr::mutate(percentage = n / sum(n))


ggplot(diamonds, aes(x = price, fill=cut))+
  geom_area(stat="bin",position="fill")+
  ggtitle("Stacked Area Graph for Comparison + Part to Whole")


ggplot(diamonds, aes(x = price, fill=cut))+
  geom_area(stat="bin",position="fill", alpha=0.6 , size=1, colour="black")+
  ggtitle("Stacked Area Graph for Comparison + Part to Whole")
```


#scale y axis to percentage

```{r}
library(scales)
ggplot(diamonds, aes(x = price, fill=cut))+
  geom_area(stat="bin",position="fill")+
  scale_y_continuous(labels = percent_format())+
  ggtitle("Stacked Area Graph for Comparison + Part to Whole")
```



## Tree Maps
### Uses: comparison,part to whole


```{r}
library(treemapify)
library(dplyr)
temp<-dplyr::summarise(group_by(diamonds,cut),total_value=sum(price), total_carat=sum(carat))


ggplot(temp, aes(area=total_value,label=cut,fill=cut))+
  geom_treemap()+ 
  geom_treemap_text(fontface = "italic", colour = "white", place = "centre",grow = FALSE)+
  ggtitle("Tree Maps for Comparison + Part to Whole")


# increasing the square font size based on the values

ggplot(temp, aes(area=total_value,label=cut,fill=cut))+
  geom_treemap()+ 
  geom_treemap_text(fontface = "italic", colour = "white", place = "centre",grow = TRUE)+
  ggtitle("Tree Maps for Comparison + Part to Whole")
```

#Alluvial Chart
# So parallel sets are useful for comparison with hierarchy. On this example 
# passengers on the titanic are stratified by demographic and survival 
 

# you remember we discussed in the first class that we have 4 variables which 3 of them are axis and a frequency value as y

```{r}
library(ggalluvial)
titanic_wide <- data.frame(Titanic)
ggplot(data = titanic_wide,
       aes(axis1 = Sex, axis2 = Class, axis3 = Age,
           y = Freq)) +
  scale_x_discrete(limits = c("Sex", "Class", "Age"), expand = c(.1, .05)) +
  xlab("Demographic") +
  geom_alluvium(aes(fill = Survived)) +
  geom_text(stat = "stratum", label.strata = TRUE)




# what if we want to show the boxed per axis/variables like a box per each class category
ggplot(data = titanic_wide,
       aes(axis1 = Sex, axis2 = Class, axis3 = Age,
           y = Freq)) +
  scale_x_discrete(limits = c("Sex", "Class", "Age"), expand = c(.1, .05)) +
  xlab("Demographic") +
  geom_alluvium(aes(fill = Survived)) +
  geom_text(stat = "stratum", label.strata = TRUE)+
  geom_stratum() 
```


## Pie Charts
### Uses: Part to whole, comparison

```{r}
# Simple Pie Chart
slices <- c(10, 12,4, 16, 8)
lbls <- c("US", "UK", "Australia", "Germany", "France")
pie(slices, labels = lbls, main="Pie Chart of Countries")
```

#Add percentage to each slice

```{r}
slices <- c(10, 12, 4, 16, 8) 
lbls <- c("US", "UK", "Australia", "Germany", "France")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, main="Pie Chart of Countries")
```
#Poolar cordinates - stacked bar chart in polar coordinates.



```{r}

# The polar coordinate system is most commonly used for pie charts 
ggplot(data = diamonds, mapping = aes(x = cut, fill = cut)) + 
    geom_bar(width = 1) + 
  coord_polar(theta = "x")
```
# So far we have seen some of the common plots which you may use in a daily manner at work/research
# the idead behind visualisation is not only how to plot or what plot to use
# Sometimes you could use a simple plot but represent it in a neat way (using better coloring, fonts, size, ...)

## How to format main, axis and legend titles


```{r}
# Main and axis title
library(ggplot2)
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

p<-ggplot(ToothGrowth, aes(x=dose, y=len)) + 
  geom_boxplot()+ ggtitle("Plot of length by dose") + 
  xlab("Dose (mg)") + ylab("Teeth length")

p
```


```{r}
# Change text of legend
p<-ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) + 
  geom_boxplot()+ ggtitle("Plot of length by dose") + 
  xlab("Dose (mg)") + 
  ylab("Teeth length") 

p+labs(fill = "Dose (mg)")

```

```{r}
p + theme(
  plot.title = element_text(color="red", size=14, face="bold.italic"),
  axis.title.x = element_text(color="blue", size=14, face="bold"),
  axis.title.y = element_text(color="#993333", size=14, face="bold")
)
```

```{r}
# Change text properties for entire plot
p+theme(text=element_text(size=16))
```



# Change legend position

```{r}
p + theme(legend.position="top")
```

```{r}
p + theme(legend.position="bottom")
```

```{r}
p + theme(legend.position = c(0.8, 0.2))
```

# Chnage legend title

```{r}
p + 
  theme(legend.title = element_text(colour="blue", size=16, face="bold"))+
  labs(fill = "Dose (mg)")
```

# Change legend text
```{r}
p + 
  theme(legend.text = element_text(colour="blue", size=16, face="bold"))+
  labs(fill = "Dose (mg)")
```

```{r}
# Change background color of legend
p + theme(legend.background = element_rect(fill="lightblue", size=0.5,linetype="solid"))
```

# Remove only the legend title using element_bank()
```{r}
p + theme(legend.title = element_blank())
```

# Remove the plot legend
```{r}
p + theme(legend.position='none')
```

```{r}
p <- ggplot(data = mtcars, 
    aes(x=mpg, y=wt, color=cyl, size=qsec, shape=factor(gear)))+
    geom_point()

p
```

# You can Select which dimension to show using legend

```{r}
p+guides(color = FALSE, size = FALSE)
```

## Changing Colors using name and numbers


```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
mtcars$cyl <- as.factor(mtcars$cyl)
```




```{r}
# box plot
ggplot(ToothGrowth, aes(x=dose, y=len)) +
  geom_boxplot( color="darkred")

# Use color names from R
```

```{r}
# box plot
ggplot(ToothGrowth, aes(x=dose, y=len)) +
  geom_boxplot(fill="burlywood", color="darkred")
```
# Use Hexcolor scheme


```{r}
# box plot
ggplot(ToothGrowth, aes(x=dose, y=len)) +
  geom_boxplot(fill="#bada55", color="#1d1916")

```

```{r}
# scatter plot
ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point(color='#3b5998')

```

```{r}
# Box plot
bp<-ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) +
  geom_boxplot()
bp
```

# Fill group colors manually

```{r}
bp+ scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))
```

# Fill group colors manually


```{r}
sp<-ggplot(mtcars, aes(x=wt, y=mpg, color=factor(cyl)))+geom_point() 

sp + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))

```
# Using palette colors from brewer package


```{r}

library(RColorBrewer)
# Box plot
bp + scale_fill_brewer(palette="Dark2")
```

```{r}
# Using grey scale
# Box plot
bp + scale_fill_grey() 
```

```{r}
# Scatter plot
sp + scale_color_grey() 
```

```{r}
# Using continuous colors
sp2<-ggplot(mtcars, aes(x=wt, y=mpg, color=qsec)) + geom_point()
sp2
```

```{r}
# Sequential color scheme
sp2+scale_color_gradient(low="blue", high="red")
```

```{r}
# Diverging color scheme
mid<-mean(mtcars$qsec)
sp2+scale_color_gradient2(midpoint=mid, low="blue", mid="white",
                     high="red", space ="Lab" )
```

# Gradient between n colors

```{r}
sp3<-ggplot(mtcars, aes(x=wt, y=mpg, color=mpg)) + geom_point()
sp3+scale_color_gradientn(colours = rainbow(2))
```


## Changing background theme

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
p <- ggplot(ToothGrowth, aes(x=dose, y=len)) + geom_boxplot()
p
```
```{r}
p + theme_gray()
```

```{r}
p + theme_bw()
```

```{r}
p + theme_light()
```

```{r}
p + theme_minimal()
```

```{r}

# Lines will disappear 
p + theme_classic()
```

```{r}
# You can remove the theme totally 
p + theme_void()
```

```{r}
p + theme_dark()
```

```{r}
ggplot(ToothGrowth, aes(x=dose, y=len, fill=supp)) +
  geom_boxplot()+
  scale_fill_manual(values=c("#fe5970", "#c2fe59"))+
  theme_dark()
```

# Custom background colors
# Change the colors of plot panel background to lightblue
# and the color of grid lines to white


```{r}

p + theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.border = element_rect(colour = "black", fill=NA, size=5),
  
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "red")
  
  )
```


# Remove panel borders and grid lines

```{r}
p + theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
```

# Hide panel borders and grid lines


```{r}
# But change axis line
p + theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))
```

## Axis Ticks

```{r}
# scatter plot
p <- ggplot(ToothGrowth, aes(x=dose, y=len)) + geom_boxplot()
p
```
# Change the appearance and the orientation angle of axis tick labels


```{r}

p + theme(axis.text.x = element_text(face="bold", color="#993333", 
                           size=14, angle=45),
          axis.text.y = element_text(face="bold", color="#993333", 
                           size=14, angle=45))
```
# Hide x an y axis tick mark labels


```{r}
p + theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank())
```

# Remove axis ticks and tick mark labels

```{r}
p + theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks = element_blank())
```

# Change the line type and color of axis lines

```{r}
p + theme( axis.line = element_line(colour = "darkblue", 
                      size = 1, linetype = "solid"))
```
# Change the order of items Change the x axis name

```{r}

p + scale_x_discrete(name ="Dose (mg)", 
                    limits=c("2","1","0.5"))
```

```{r}
# Change ticks name
p + scale_x_discrete(breaks=c("0.5","1","2"),
        labels=c("Dose 0.5", "Dose 1", "Dose 2"))
```

# Choose which items to display

```{r}
p + scale_x_discrete(limits=c("0.5", "2"))
```

# Changing axis with continuous variables

```{r}
# lets make a refrence plot
sp<-ggplot(cars, aes(x = speed, y = dist)) +
  geom_point()

sp
```

# Change x and y axis labels, and limits

```{r}
sp + scale_x_continuous(name="Speed of cars", limits=c(0, 30)) +
  scale_y_continuous(name="Stopping distance", limits=c(0, 150))
```

# Set tick marks on y axis a tick mark is shown on every 5


```{r}
p + scale_y_continuous(breaks=seq(0,40,5))
```



# Tick marks can be spaced randomly

```{r}
p + scale_y_continuous(breaks=c(5,7.5, 20, 25))
```

# Remove tick mark labels and gridlines

```{r}
p + scale_y_continuous(breaks=NULL)
```

# Multigrid - why its good?

```{r}
p <- ggplot(mpg, aes(displ, cty)) + geom_point()

# Use vars() to supply variables from the dataset:
p + facet_grid(rows = vars(drv))
```

```{r}
p <- ggplot(mpg, aes(displ, cty)) + geom_point()

# Use vars() to supply variables from the dataset:
p + facet_grid(cols  = vars(drv))
```


```{r}
p + facet_grid(vars(drv), vars(cyl))
```

```{r}
p + facet_grid(. ~ cyl)
```

```{r}
p + facet_grid(cyl~.)
```

```{r}
p + facet_grid(drv ~ cyl)
```

# Separataing for each factor (cyl)

```{r}
mt <- ggplot(mtcars, aes(mpg, wt, colour = factor(cyl))) +
  geom_point()

mt + facet_grid(. ~ cyl, scales = "free")
```


# Separataing for multiple conditions (we will get a plot per combination)


```{r}
mg <- ggplot(mtcars, aes(x = mpg, y = wt)) + geom_point()
mg + facet_grid(carb + am ~ gear)
```




# Adding the label to eahc box on both col and rows (here cyl: and drv:) are labels

```{r}
p + facet_grid(drv ~ cyl, labeller = label_both)
```

```{r}
mg + facet_grid(
  vs + am ~ gear,
  labeller = labeller(.rows = label_both, .cols = label_value)
)


```


# edit the headers 

```{r}
p + facet_grid(drv ~ cyl)+
  theme(strip.background = element_rect(colour="black", fill="white", size=1.5, linetype="solid"))
```

```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point(color="#ffd23d") + 
  facet_grid(drv ~ cyl)+
  theme_dark(base_size = 18)
```


```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point(color="#2f6276") + 
  facet_grid(drv ~ cyl)+
  theme_bw(base_size = 18)
```

```{r}
ggplot(mpg, aes(displ, cty)) + 
  geom_point(color="#008f95") + 
  facet_grid(drv ~ cyl)+
  theme_minimal(base_size = 18)
```

```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point(color="#2f6276") + 
  facet_grid(drv ~ cyl)+
  theme_classic()
```

# save a plot

```{r}
ggplot(mpg, aes(displ, cty)) + 
  geom_point(color="#2f6276") + 
  facet_grid(drv ~ cyl)+
  theme_bw(base_size = 18)
ggsave("output.png", device="png")
```








